---
title: "Group Exercise"
author: Group 7
date: Nov 18, 2023
output: 
    html_document:
        # latex_engine: xelatex
        toc: true
        toc_depth: 2
---

# Data description

```{r}
decathlon_data <- read.csv(
  "decathlon.csv", 
  row.names=1, 
  sep=";"
)

#Variables and individuals
variables <- colnames(decathlon_data[, 1:10])
individuals <- rownames(decathlon_data[, 1:10])
print(paste("Variables:", paste(variables, collapse = ", ")))
print(paste("Individuals:", paste(individuals, collapse = ", ")))
```

# Purpose of PCA

=> PCA aims to identify the most influential events contributing to athletes overall success. By extracting key patterns and relationships among the 10 events, PCA facilitates dimensionality reduction, revealing which competitions significantly impact final rankings. This allows for a concise representation of athletes performance profiles, aiding in visualization, pattern recognition, and the identification of critical features shaping decathlon outcomes overall success.

# Scale

```{r}
scaled_data <- scale(decathlon_data[,1:10])

scaled_df <- data.frame(scaled_data, decathlon_data[, (ncol(decathlon_data)-1):(ncol(decathlon_data))])

print(scaled_df)
```

This process involves transforming the data so that it has a mean of 0 and a standard deviation of 1. This is to ensures that all variables are on a comparable scale, preventing one variable from dominating the analysis due to its larger magnitude.
Data reduction through standardization is crucial in the decathlon analysis because it enables fair and meaningful comparisons between events, ensures the sensitivity of PCA, and enhances the interpretability and visual analysis of athlete performance profiles.

# Calculate the correlation matrix

```{r}
cor_matrix <- round(cor(scaled_data), 2)
print(cor_matrix)
```

# Applying PCA on data

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(ggplot2)
library(FactoMineR)
library(factoextra)
library(FactoMineR)
library(Factoshiny)
library(httpgd)
library(reshape2)
```

## Data mining

```{r}
df <- read.csv(
  "decathlon.csv", 
  row.names=1, 
  sep=";"
)

# Prepare data

df$Competition <- as.factor(df$Competition)

# Normalize data

df.norm <- data.frame(scale(df[,1:10]), df[,11:13])

cor(df[, c(1:12)])

pairs(df[, c(1:12)])
```

## Apply PCA

```{r}
res.pca <- PCA(
  df.norm, 
  quanti.sup = c("Rang", "Points"), 
  quali.sup = "Competition",
  graph = F
)
```

## Eigenvalues

### Eigenevalues Table

```{r}
  res.pca$eig[1:5, ]
```

### Eigenvalues Plot

```{r}
  fviz_eig(res.pca)
```

First four components have a eigenvalue value of greater than 1.

However, the first two dimension comprising more than 50% of the total variability of the cloud of individuals (or variables) is represented by the plane, it is adequate to expresses a significant structure in the data.

## Individuals

### General Plot

```{r}
plot.PCA(
  res.pca, 
  axes = c(1,2),
  choix = "ind", 
  invisible = c("quali", "var"),
  select = "cos2 0.4",
  unselect = 0.8
)
```

-   Observation:

Highly correlated to dim1 than dim2: MAYER, WIESIOLEK, LELIEVRE

Highly correlated to dim2 than dim1: Kaul, Suarez, Zhuk, GADO

### By Points

```{r}
  plot.PCA(
    res.pca, 
    axes = c(1,2),
    choix = "ind", 
    habillage = "Points", 
    invisible = c("quali", "var"),
    select = "cos2 0.4",
    unselect = 0.8
  )
```

-   Observation:

Individuals positively correlated to dim 1 has the highest scores, and ones with negative correlation to dim 1 has lowest scores.

### By Competition

```{r}
  plot.PCA(
    res.pca, 
    axes = c(1,2),
    choix = "ind", 
    habillage = "Competition", 
    invisible = c("quali", "var"),
    select = "cos2 0.4",
    unselect = 0.8
  )
```

## Variables

### General Plot

```{r}
  plot(res.pca,
     axes = c(1, 2),
     choix="var",
     hab="none",
     invisible = c("quali", "ind"),
     select="cos2 0.4",
     unselect=0.8
  )

```

# Interpretating PCA results

## Main axes of PCA

```{r}
  res.km <- kmeans(res.pca$var$coord, centers = 3, nstart=25)
  grp <- as.factor(res.km$cluster)
  plot.PCA(res.pca, choix="var", col.var=grp, select="cos2 0.3")
  # fviz_pca_var(res.pca, axes=c(3,4), repel=T, col.var=grp)
```

### 1st dimension

From the plot of correlation, we can see that most running subjects (**X100m**, **X400m** and **X110.haies**) contribute significantly to the first dimension of PCA. Those subjects have high cos2 (higher than 0.5) and strong negative correlation to this dimension. Another observation is that **Range** (quantitative supplementary variable) also have high cos2 and strong negative correlation to dimension 1, whereas **Points** (another quantitative supplementary variable) demonstrates nearly perfect positive correlation with the 1st dimension and cos2 is very close to 1.0 (0.97). We can interpret that, the less time an athelete requires to complete a run, the more likely they are to gain higher points. Furthermore, Longueur have high cos2 in the main axes and almost lies in the opposite side of running subjects, indicating that an atheletes who are good at running also good at Longeur.

Let demonstrate the above observation by plotting some individual plots

```{r}
  plot.PCA(res.pca, habillage ="X100m", select="cos2 0.4")
  plot.PCA(res.pca, habillage ="X110m.haies", select="cos2 0.4")
  plot.PCA(res.pca, habillage ="Longueur", select="cos2 0.4")
  plot.PCA(res.pca, habillage ="Points", select="cos2 0.4")
```

As can be seen from these plots, from left to right of dimension 1, the time needed by an athelete to complete **100m run** or **110m haies run** *decreases*. In **Longueur** plot, athelete in the right of dimension 1 exhibit better performance (higher distance) comparing to those on the left. The point gained by athelte on the right side of dimension 1 is higher than those on the left.

For example, MAYER, Clayer and Warner both have good achievement in 100m Running, 110m Haies Running and Longueur. They also gained high points and top range, as shown below (comparing to Schawarzl, Kaul that in the middle of the axes and NILSSON, BARAN who are in the left side).

```{r}
  df[
    c('MAYER', 'Clay', 'Warner', 'Schwarzl', 'Kaul', 'NILSSON', 'BARAN'), 
    c('X100m', 'X110m.haies', 'Longueur', 'Points', 'Rang')
  ]
```

To conclusion, the first dimension of PCA represent performance of atheletes on **running subjects** and **Longueur**. This dimension also represent the **point** and the **range** of an athelete.

### 2nd dimension

The dimension 2 of PCA, however, does not well represent any subject. Some subjects that have *relatively high* cos2 to this dimension are mainly throwing sports such as **Javelot** (cos2=0.34), **Disque** (cos2=0.22) and **Poids** (cos2=0.17). Some sports also contribute fairly much to this dimension are X100m (cos2=0.29) and Longueur (cos2=0.24). Those subjects also have **high cos2 with the first dimension**. Because of that, atheletes that have high performance on those subject are positioned not only on the top of dimension 2, but also on the right of dimension 1, as can be seen from the following individual plots:

```{r}
  plot.PCA(res.pca, habillage ="Javelot", select="cos2 0.4")
  plot.PCA(res.pca, habillage ="Disque", select="cos2 0.4")
  plot.PCA(res.pca, habillage ="Poids", select="cos2 0.4")
```

Let choose some individuals to see their performance:\

-   Kaul, Suarez, Zhuk: Top of dimension 2.
-   Sebrle, Clay, MAYER: Right of dimension 1.
-   WIESIOLEK: Left of dimension 1, In the middle of dimension 2.
-   Warners: In the middle of dimension 1, Bottom of dimension 2.

```{r}
  df[
    c("Kaul", "Suarez", "Zhuk", "Sebrle", "Clay", "MAYER", "WIESIOLEK", "Warners"), 
    c("Javelot", "Disque", "Poids", "Longueur", "Points", "Rang")]
```

As we can see, Kaul, Suarez and Zhuk have better performance in **Javelot** and **Disque** than WIESIOLEK and Warners, who are in lower positions in dimension 2 and low position in dimension 1. Warners, despite being at the low position of dimension 2, still have similar performance in Poids to Kaul, Suarez and Zhuk since he is further to the right in dimension 1 than those three. However, Kaul, Suarez and Zhuk are not as good in **Disque** and **Poids** as Sebrle, Clay and MAYER, who are in lower positions in dimension 2 but on the right side of dimension 1.

We can conclude that, dimension 2 can somehow represents athelete performance in **throwing sports** and **Longueur** as people in the top position in this dimension tend to achieve *above-average performance* in these sports.

### 3rd dimension

Using `dimdesc`, we can get 4 variables that are well represented on dimension 3,

```{r}
  dimdesc(res.pca)$`Dim.3`
```

which are: **X1500m**, **Disque**, **Perche** and **X400m**. The variable PCA plot below can also demonstrate it.

```{r}
  plot.PCA(
    res.pca, 
    choix="var", 
    axes=c(2,3), 
    invisible = "quanti.sup", 
    select="cos2 0.2"
  )
```

Let plot individual plot using those variable as habillage:

```{r}
  plot.PCA(res.pca, axes=c(2,3), habillage ="X1500m", select="cos2 0.4")
  plot.PCA(res.pca, axes=c(2,3), habillage ="Disque", select="cos2 0.4")
  plot.PCA(res.pca, axes=c(2,3), habillage ="Perche", select="cos2 0.4")
  plot.PCA(res.pca, axes=c(2,3), habillage ="X400m", select="cos2 0.4")
  
```

We can clearly see that athelete positioned at the top of dimension 3 have better performance on X1500m, X400m, Disque, Perche. We can validate this by considering 4 atheletes:\
- Duckworth, Ziemek: Top position of dimension 3 - Cairoli, Bourrada: Bottom position of dimension 3

```{r}
  df[
      c("Duckworth", "Ziemek", "Cairoli", "Bourrada"), 
      c("X1500m", "X400m", "Disque", "Perche")
  ]
```

As we can see, Duckworth and Ziemek outperform Cairoli and Bourrada on almost listed sports, except that Duckworth is worse than Bourrada in **Disque**. This is due to cos2 of **Disque** on dimension 2 and dimension 3 are similar (0.22 and 0.19) and athelete performance on this sport is positive correlated on both of 2 dimensions, which means athelete are on the right side of dimension 2 or on the top of dimension 3 tend to have higher performance on Disque. As we can see in individual plot, although Duckworth lies higher than Bourrada on dimension 3, he lies further to the left in dimension 2 than Bourrada. This can explain why Duckworth performance on Disque are not so different than that of Bourrada.

## In what activities do decathlon winners excel? Is there a case where any competition has the least impact on performance (Points) in the decathlon?

```{r}
plot(res.pca,
     axes = c(1, 2),
     choix = c("var"),
     col.hab = "Points",
     invisible = c("ind"),
     select=NULL,
     unselect=0.8
     )
```

* We see that on the correlation circle graph there is an additional variable “Points” that is almost aligned with the first dimension. This means that decathlon winners performed well in sports that were highly correlated with the first principal component.

```{r}
res.pca$var$coord
```

* In addition, we see that the sports most closely associated with the first principal component (with coord values near 1 or -1) include: Longueur, Disque, Poids, X100m, X400m, X110m.haies. Thus, decathlon winners excel in activities that are strongly correlated with the first principal component aforementioned.

* Similarly, sports that have little impact on performance (Points) will have a low correlation with the first axis, including: Perche and X1500.

## Compare achievements in 4 competitions. What can you say about the average performance in this competition?

```{r}
plot(res.pca,
     axes = c(1, 2),
     choix="ind",
     hab="Competition",
     invisible = c("var","quanti"),
     select=NULL,
     unselect=0.8
     )

plotellipses(res.pca, label="none")
```

At the 0.95 confidence level, the confidence ellipses split to the right (JO04 and JO16) and left (Decastar18 and CE18). The above observation may indicate that decathlon competitions and scoring systems favor a number of sports closely related to the first principal component, in which the majority of athletes have better performance in the JO04 and JO08. 

Additionally, between JO04 and JO16, there are overlapping confidence ellipses, so they are almost identical, but there are differences between these two competitions with Decastar18 and CE18.

There is not too much overlap between Decastar18 and CE18, so there is still a difference between the two exams.

```{r}
plotellipses(res.pca, label="none", level = .5)
```

At the 0.50 confidence level, the confidence intervals for JO04 and JO18 still overlap, while those for Decastar18 and CE18 are separated.



```{r}
library(FactoMineR)
res.PCA <- PCA(decathlon, quali.sup = 13, quanti.sup = c(11,12), graph = TRUE, ncp = 8)
summary(res.PCA)
```

## Who are the top 5 athletes who performed best on the main principle component?

From the PCA above, we can deduce the top 5 outstanding athletes on the main plane are Sebrle, Clay, Karpov, Macey, and Warners because all five of them have higher overall scores in dimensions 1, 2, and 3 compared to the others.

## Who are the athletes who contributed the most to building the first principle component of the ACP? And in the second? What does that significant contribution mean?

The athlete who contributes the most to the construction of the first principal axis of PCA is MAYER.
The athlete who contributes the most to the construction of the second principal axis of PCA is NILSSON.
This suggests that MAYER has a higher score in more events compared to NILSSON, who has a high score in fewer events.

## From the correlation circle, what can be said about the following correlations:

From the correlation circle:
+ 100m hurdles - 400m hurdles: they have a very small angle between them -> The athletes' scores in these two events have a similar difference.
+ 100m - Long jump: they are opposite to each other by 180 degrees -> These two events represent two unrelated fields, but the angle with the point representing the scores in both events is very small.
+ Disque - Poids: they are close to each other in terms of angles towards the points -> These two events have a high correlation in dimension 1.
+ Disque - Long jump: they are on the same side of the point but symmetrically across dimension 1 by approximately 90 degrees -> These two events are not correlated.
+ Longueur - 1500m: they are opposite to each other by 180 degrees -> These two events represent two unrelated fields, but the scores in both events are very high.